# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
First we read given .csv file and convert column
```{r}
steps_data_raw <- read.csv("activity.csv")  
steps_data_raw$date <- as.Date(steps_data_raw$date)
steps_data_daily <- aggregate(steps~date, steps_data_raw, sum)
steps_data_mean_interval <- aggregate(steps~interval, steps_data_raw, mean)
```

## What is mean total number of steps taken per day?
```{r fig.width=8, fig.height=8, message=FALSE}
hist(aggregate(steps~date, steps_data_raw, sum)$steps, col="green")
```

```{r}
steps_mean <- mean(steps_data_daily$steps)
steps_median <- median(steps_data_daily$steps)
```

The average of steps taken each day is *`r steps_mean`*  
The median of steps taken each day is *`r steps_median`*

## What is the average daily activity pattern?
```{r fig.width=8, fig.height=8, message=FALSE}
plot(x=steps_data_mean_interval$interval, y=steps_data_mean_interval$steps, type="l")
```

To find 5-minute interval containing maximum average number of steps, let's first find out the maximum average value of steps. Then we can find value of `interval` variable corresponding to max steps value in data set.
```{r}
max_steps_value <- max(steps_data_mean_interval$steps)
max_steps_interval <- steps_data_mean_interval[steps_data_mean_interval$steps == max_steps_value, ]$interval
```

Index of 5-minute interval, which contains maximum number of steps average across whole dataset, is *`r max_steps_interval`*

## Imputing missing values
```{r}
num_missing_values <- sum(!complete.cases(steps_data_raw))
```
Total number of rows containing missing values is *`r num_missing_values`*

For each date missing values are going to be replaced by mean steps value for that date. If all steps values are missing during particular date, they will be replaced with 0
```{r}
#steps_data_clean is the frame containing approximated na values
steps_data_clean <- steps_data_raw
steps_data_clean$date <- as.Date(steps_data_clean$date)

steps_data_daily_mean <- aggregate(steps~date, steps_data_raw, mean)

for(date in unique(steps_data_raw$date))
{
    steps_data_raw_date <- steps_data_raw[steps_data_raw$date == date, ]
    #if we have na values on this date
    if(sum(is.na(steps_data_raw_date$steps)) > 0)
    {
        #if we don't have mean value for this date, replace na's with 0
        if(nrow(steps_data_daily_mean[steps_data_daily_mean$date == date,])==0)
        {
            steps_data_clean[steps_data_raw$date == date & is.na(steps_data_raw$steps), ]$steps <- 0
        }
        else
        {
            #if we do have a mean steps value for this date, replace all NAs with this mean steps value
            steps_data_clean[steps_data_raw$date == date & is.na(steps_data_raw$steps), ]$steps <- steps_data_daily_mean[steps_data_daily_mean$date == date, 'steps']
        }
    }
}

steps_data_daily_clean <- aggregate(steps~date, steps_data_clean, sum)
```

```{r fig.width=8, fig.height=8, message=FALSE}
hist(aggregate(steps~date, steps_data_clean, sum)$steps, col="green")
```

```{r}
steps_mean_clean <- mean(steps_data_daily_clean$steps)
steps_median_clean <- median(steps_data_daily_clean$steps)
```

The average of steps taken each day in clean frame is *`r steps_mean_clean`*  
The median of steps taken each day in clean frame is *`r steps_median_clean`*

As we can see, mean and median are shifted to 0. Also, histogram shows that there are more 0 values than in frame containing NAs. This indicates that most of NAs were replaced by 0 values.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
steps_data_clean$weekend <- ifelse(weekdays(steps_data_clean$date) %in% c("воскресенье", "суббота"), "weekend", "weekday")
```

```{r fig.width=8, fig.height=8, message=FALSE}
par(mfrow = c(2,1))
steps_data_weekday <- aggregate(steps~interval, subset(steps_data_clean, weekend == "weekday"), mean)
steps_data_weekend <- aggregate(steps~interval, subset(steps_data_clean, weekend == "weekend"), mean)
with(steps_data_weekday, plot(interval,steps, main="Weekday", type="l"))
with(steps_data_weekend, plot(interval,steps, main="Weekend", type="l"))
```